The \dealii Library, Version 9.1

### 2019-05-14 9.1/paper.tex
List GPU support

The major changes of this release are:
\begin{itemize}
\item Improved support for automatic and symbolic differentiation;
\item Full support for $hp$ adaptivity in parallel computations;
\item Interfaces to the HDF5 file format and libraries.
\item Four new tutorial programs step-61, step-62, step-63, step-64,
  as well as one new code gallery program.
\end{itemize}
These will all be discussed in more detail in the
following section. There are a number of other noteworthy changes in this release:
\begin{itemize}
\item A number of bug fixes and performance improvements to the
  matrix-free framework and related geometric multigrid solvers. In
  particular, the implementation of the Chebyshev iteration, an often
  used smoother in the matrix-free context, has been revised to reduce
  vector accesses. This speeds up matrix-free
  multigrid solvers by up to 10-15\% on geometries with affine
  (parallelogram and parallelpiped) cells, and up to
  5\% on geometries with cells bounded by curved edges and faces.
\marginpar{Luca: Please expand}
\item ParsedConvergenceTable.
\item The \texttt{FE\_BernardiRaugel} class implements the
  non-standard Bernardi-Raugel (BR) element that can be used as one part of
  a stable velocity/pressure pair for the Stokes equation
  \cite{BR85}. The BR element can be seen as either an enriched
  version of the $Q_1^d$ element with added bubble functions on each
  edge (in 2d) or face (in 3d), or as a reduced version of the $Q_2^d$
  element. It addresses the fact that the $Q_1^d\times Q_0$
  combination is not inf-sup stable (requiring a larger velocity
  space), and that the $Q_2^d\times Q_1$ combination is stable but
  sub-optimal since the velocity space is too large relative to the
  pressure space to provide additional accuracy commensurate with the
  cost of the large number of velocity unknowns. The BR space is
  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.

@@

@@

The original \texttt{\dealii} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019,KronbichlerLjungkvist2019};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
  \item For partition-of-unity (PUM) and enrichment methods of the
    finite element space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012,KronbichlerKormann2019};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For Boundary Elements Computations:
   \cite{GiulianiMolaHeltai-2018-a};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16}.
\end{itemize}

>>>>

The major changes of this release are:
\begin{itemize}
\item Improved support for automatic and symbolic differentiation;
\item Full support for $hp$ adaptivity in parallel computations;
\item Interfaces to the HDF5 file format and libraries.
\item Significantly extended GPU support.
\item Four new tutorial programs step-61, step-62, step-63, step-64,
  as well as one new code gallery program.
\end{itemize}
These will all be discussed in more detail in the
following section. There are a number of other noteworthy changes in this release:
\begin{itemize}
\item A number of bug fixes and performance improvements to the
  matrix-free framework and related geometric multigrid solvers. In
  particular, the implementation of the Chebyshev iteration, an often
  used smoother in the matrix-free context, has been revised to reduce
  vector accesses. This speeds up matrix-free
  multigrid solvers by up to 10-15\% on geometries with affine
  (parallelogram and parallelpiped) cells, and up to
  5\% on geometries with cells bounded by curved edges and faces.
\item Various variants of geometric multigrid solvers and matrix-free
  implementations were run on up to 304,128 cores during the acceptance phase of
  the SuperMUC-NG supercomputer, verifying the scalability of our
  implementations. Some performance bottlenecks on distributed triangulation
  with more than 100k MPI ranks have been fixed.
\marginpar{Luca: Please expand}
\item ParsedConvergenceTable.
\item The \texttt{FE\_BernardiRaugel} class implements the
  non-standard Bernardi-Raugel (BR) element that can be used as one part of
  a stable velocity/pressure pair for the Stokes equation
  \cite{BR85}. The BR element can be seen as either an enriched
  version of the $Q_1^d$ element with added bubble functions on each
  edge (in 2d) or face (in 3d), or as a reduced version of the $Q_2^d$
  element. It addresses the fact that the $Q_1^d\times Q_0$
  combination is not inf-sup stable (requiring a larger velocity
  space), and that the $Q_2^d\times Q_1$ combination is stable but
  sub-optimal since the velocity space is too large relative to the
  pressure space to provide additional accuracy commensurate with the
  cost of the large number of velocity unknowns. The BR space is
  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GPU support via CUDA}

The GPU support was significantly extended for the current release. On the one hand, the MPI-parallel vector class
\texttt{LinearAlgebra::distributed::Vector} has gained a second template
argument \texttt{MemorySpace} which can either be \texttt{Host} or
\texttt{CUDA}. In the latter case, the data resides in the GPU memory. Also,
the ghost exchange can be performed completely via CUDA in case a compatible
MPI library is found (as is usual on large GPU-based supercomputers).

Furthermore, a new tutorial program \texttt{step-64} has been added that demonstrates the usage of matrix-free methods on Nvidia GPUs. GPUs are advantageous for these kind of operations because of their superior hardware characteristics, in particular a higher memory bandwidth than server CPUs within a given power envelope. The matrix-free GPU components integrated in \dealii have been analyzed against CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric multigrid solvers is discussed.

@@

The original \texttt{\dealii} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
  \item For partition-of-unity (PUM) and enrichment methods of the
    finite element space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012,KronbichlerKormann2019};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For Boundary Elements Computations:
   \cite{GiulianiMolaHeltai-2018-a};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16}.
\end{itemize}
